---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files

# packages for using custom fonts
# library(systemfonts)
# library(textshaping)
library(ragg)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

Saving using custom fonts with ragg

```{r}
ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  labs(title = "22222222222222222") +
  theme(text = element_text(family = "Roboto"),
        plot.title = element_text(size = 50))

ggsave("filename.png", device = ragg::agg_png())
```


## Colour palettes

## Remember it is an iterative process

## Import data

```{r}
production <- read_csv("20211012/production.csv") %>% 
  janitor::clean_names() %>% 
  rename(seafood_prod = new_food_balances_fish_seafood_2960_production_5511_1000_tonnes)

consumption_pc <- read_csv("20211012/consumption_pc.csv") %>% 
  janitor::clean_names() %>% 
  rename(seafood_con_pc = fish_seafood_food_supply_quantity_kg_capita_yr_fao_2020)

population <- read_csv("20211012/population.csv") %>% 
  janitor::clean_names() %>% 
  rename(population = population_historical_estimates_and_future_projections)

library(sf)
countries <- read_sf("../../geospatial_data/vector/countries/ne_10m_admin_0_countries/ne_10m_admin_0_countries.shp")

ggplot() +
  geom_sf(data = countries)
  
```

## Clean data

May be use these shape files for country boundaries

https://www.naturalearthdata.com/downloads/10m-cultural-vectors/10m-admin-0-details/

```{r}
prod_2017 <- production %>% 
  filter(year == 2017)

con_pc_2017 <- consumption_pc %>% 
  filter(year == 2017)

pop_2017 <- population %>% 
  filter(year == 2017)

net_2017 <- prod_2017 %>% 
  full_join(con_pc_2017) %>% 
  left_join(pop_2017) %>% 
  na.omit() %>% 
  mutate(seafood_con = seafood_con_pc * population / 1000,
         net = seafood_prod - seafood_con)

View(net_2017)
View(countries)

library(colorspace)

plotting_df <- countries %>% 
  select(ISO_A3_EH) %>% 
  rename(code = ISO_A3_EH) %>% 
  left_join(net_2017) %>% 
  mutate(code = na_if(code, "-99")) %>% 
  filter(!is.na(code))

plotting_missing <- plotting_df %>% 
  filter(is.na(net)) %>% 
  filter(code != "ATA")

plotting_data_available <- plotting_df %>% 
  filter(!is.na(net)) %>% 
  mutate(net_binned = cut(net/1e6, breaks = c(-3, -0.1, 0.1, 9)))
  
  ggplot() +
  geom_sf(data = plotting_missing, size = 0.1, fill = "grey90", colour = "white") +
  geom_sf(data = plotting_data_available,
          mapping = aes(fill = net_binned), size = 0.1, colour = "white") +
  #scale_fill_distiller(palette = "RdBu")
  # colorspace::scale_fill_continuous_divergingx(mid = 0, palette = "Tropic", 
  #                                              rev = TRUE, limits = c(-2.5, 10)) +
  scale_fill_manual(values = c("#BD6F97", "#E6E7E4", "#6F97BD")) +
  theme_void() +
  theme(legend.position = "top",
        plot.background = element_rect(fill = "grey97"))


ggsave("20211012/map.svg")
  
  # summarise(total_prod = sum(seafood_prod),
  #           total_con = sum(seafood_con))


plotting_data_available %>% 
  st_drop_geometry() %>% 
  filter(net_binned != "(-0.1,0.1]") %>% 
  mutate(entity = fct_reorder(entity, net)) %>% 
  distinct() %>% 
  
  ggplot() +
  geom_col(aes(x = net, y = entity, fill = net_binned)) +
  scale_fill_manual(values = c("#BD6F97", "#6F97BD")) +
  theme_cjmr_explanatory()
  
```

## Explore the data

```{r}

```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-------+
| To think about                                                                    | Notes |
+===================================================================================+=======+
| Can you think of a group of people who might view the visualisations?             |       |
+-----------------------------------------------------------------------------------+-------+
| Can you think of a specific person who might view the visualizations?             |       |
+-----------------------------------------------------------------------------------+-------+
| What do want the audience to know or be able to after viewing the visualizations? |       |
+-----------------------------------------------------------------------------------+-------+
| What expectations might the audience have about visualisations of this data?      |       |
+-----------------------------------------------------------------------------------+-------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Question to consider                                | Tips                                                         | Notes    |
+=====================================================+==============================================================+==========+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                  |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Is the **context** explained?                       | Titles, subtitles                                            |          |
|                                                     |                                                              |          |
|                                                     | Annotate                                                     |          |
|                                                     |                                                              |          |
|                                                     | Use colour to highlight                                      |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*           |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Do the **aesthetics** align with the topic/message? | Serious topic -\> dark colours, straight-forward chart types |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+

## Remember it is an iterative process

