---
title: "Tidy Tueaday: Bird Baths"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files
library(patchwork)  # for composing layouts of multiple plots

# packages for using custom fonts
library(systemfonts)
library(textshaping)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

## Colour palettes

## Remember it is an iterative process

## Import data

```{r}
bird_baths <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-31/bird_baths.csv')
```

## Clean data

```{r}
bird_baths %>% 
  visdat::vis_miss()
```

## Explore the data

```{r}
bird_baths %>% 
  skimr::skim()
```

```{r}
p <- ggplot(bird_baths, aes(bird_count)) +
  geom_histogram() +
  scale_y_log10() +
  facet_wrap(~survey_year)

p 

p + scale_x_continuous(limits = c(0,10))


ggplot(bird_baths, aes(factor(survey_year), bird_count)) +
  geom_boxplot()
```
```{r}
bird_baths_clean <- bird_baths %>% 
  na.omit() %>% 
  filter(bird_count > 0) 

counts_by_bioregion <- bird_baths_clean %>% 
  count(bioregions, bird_type)
  
counts_by_bioregion
```

```{r}
bird_baths_clean %>% 
  count(survey_year, bird_type) %>% 
  
  ggplot(aes(fct_reorder(bird_type, n), n)) +
  geom_col(stat = "count") +
  coord_flip() +
  facet_wrap(~survey_year)

```
```{r}
seasonal_counts <- bird_baths_clean %>% 
  count(survey_year, urban_rural, bird_type) %>% 
  pivot_wider(names_from = survey_year, values_from = n, values_fill = 0) %>% 
  mutate(change_in_obs = `2015` - `2014`) %>% 
  #filter(abs(change_in_obs) >= 10) %>% 
  arrange(bird_type, change_in_obs) %>% 
  mutate(direction_of_change = if_else(
    change_in_obs > 0, "rise", "fall"
  ))


trend_types <- seasonal_counts %>% 
  select(-direction_of_change, -`2014`, -`2015`) %>% 
  pivot_wider(names_from = urban_rural, values_from = change_in_obs, values_fill = 0) %>%
  mutate(trend_type = case_when(
    (Rural >= 0 & Urban >= 0) ~ "rising_both",
     (Rural < 0 & Urban < 0) ~ "falling_both",
     (Rural >= 0 & Urban < 0) ~ "rising_r_falling_u",
     (Rural < 0 & Urban >= 0) ~ "rising_u_falling_r",
     TRUE ~ "error")
  ) %>% 
  filter((abs(Rural) >= 5) & abs(Urban) >= 5) %>% 
  select(bird_type, trend_type)


plotting_df <- seasonal_counts %>% 
  select(-direction_of_change) %>% 
  inner_join(trend_types) %>% 
  #select(-change_in_obs) %>% 
  pivot_longer(cols = c(`2014`,`2015`), names_to = "year", values_to = "n") %>% 
  #filter(bird_type == "Australian Magpie") 
  {.}

urban_df <- plotting_df %>% 
  filter(urban_rural == "Urban")

rural_df <- plotting_df %>% 
  filter(urban_rural == "Rural")

urban_line_df <-  urban_df %>% 
  pivot_wider(names_from = year, values_from = n)

rural_line_df <-  rural_df %>% 
  pivot_wider(names_from = year, values_from = n)


  ggplot() +
  geom_point(data = urban_df, position = position_nudge(x = 0, y = 0.1),
             mapping = aes(n, bird_type, colour = year, shape = urban_rural)) +
  geom_segment(data = urban_line_df, position = position_nudge(x = 0, y = 0.1),
               mapping = aes(y = bird_type, yend = bird_type,
                             x = `2014`, xend = `2015`)) + 
    
  geom_point(data = rural_df, position = position_nudge(x = 0, y = -0.1),
             mapping = aes(n, bird_type, colour = year, shape = urban_rural)) +
    
    geom_segment(data = rural_line_df, position = position_nudge(x = 0, y = -0.1),
               mapping = aes(y = bird_type, yend = bird_type,
                             x = `2014`, xend = `2015`)) +
    facet_wrap(~trend_type, scales = "free_y") +
    theme_cjmr_explanatory()
  
  # ggplot(aes(n, bird_type, colour = year)) +
  # geom_point() +
  # facet_wrap(~trend_type)
  
  
trend_types %>% 
  count(trend_type)
```


## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-------+
| To think about                                                                    | Notes |
+===================================================================================+=======+
| Can you think of a group of people who might view the visualisations?             |       |
+-----------------------------------------------------------------------------------+-------+
| Can you think of a specific person who might view the visualizations?             |       |
+-----------------------------------------------------------------------------------+-------+
| What do want the audience to know or be able to after viewing the visualizations? |       |
+-----------------------------------------------------------------------------------+-------+
| What expectations might the audience have about visualisations of this data?      |       |
+-----------------------------------------------------------------------------------+-------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Question to consider                                | Tips                                                         | Notes |
+=====================================================+==============================================================+=======+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                  |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Is the **context** explained?                       | Titles, subtitles                                            |       |
|                                                     |                                                              |       |
|                                                     | Annotate                                                     |       |
|                                                     |                                                              |       |
|                                                     | Use colour to highlight                                      |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*           |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Do the **aesthetics** align with the topic/message? | Serious topic -\> dark colours, straight-forward chart types |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+

## Remember it is an iterative process
