---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files
library(patchwork)  # for composing layouts of multiple plots

# packages for using custom fonts
library(systemfonts)
library(textshaping)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

## Colour palettes

## Remember it is an iterative process

## Import data

From the following article - https://www.nature.com/articles/sdata201419

```{r}
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')
```

## Clean data

```{r}
lemurs %>% 
  visdat::vis_miss(warn_large_data = FALSE)

lemurs_focus <- lemurs %>% 
  select(dlc_id, sex, dob, estimated_dob, birth_type, dod, age_last_verified_y, age_category, age_max_live_or_dead_y, age_last_verified_y) 

lemurs_focus%>% 
  visdat::vis_miss(warn_large_data = FALSE)
```

## Explore the data

```{r}
ggplot(lemurs_focus, aes(age_max_live_or_dead_y)) +
  geom_histogram()


lemurs_focus %>% 
  filter(is.na(dod)) %>% 
  ggplot(aes(age_max_live_or_dead_y)) +
  geom_histogram()

lemurs_focus %>% 
  filter(is.na(dod)) %>% 
  ggplot(aes(dob)) +
  geom_histogram()

lemurs_focus %>% 
  filter(is.na(dod)) %>%  
  mutate(date_last_verified = 365 * age_last_verified_y + dob) %>% 
    ggplot(aes(date_last_verified)) +
  geom_histogram()

```

```{r}
# work out dates lemurs went missing
# assume they are missing or dead if they haven't had age verified in last two years

lemurs_focus_1 <- lemurs_focus %>% 
  # calculate date last verified
  mutate(date_last_verified = 365 * age_last_verified_y + dob)

max(lemurs_focus_1$date_last_verified, na.rm = TRUE)
min(lemurs_focus_1$date_last_verified, na.rm = TRUE)

cutoff_date <- ymd("2017-01-04")

ggplot(lemurs_focus_1, aes(date_last_verified)) +
  geom_histogram()

# lemurs_focus_2 <- lemurs_focus_1 %>% 
#   
#   # calculate date missing or dead
#   mutate(date_missing_dead = case_when(
#     ((is.na(dod)) & (date_last_verified < cutoff_date)) ~ date_last_verified,
#     TRUE ~ dod
#   ))

lemurs_focus_2 <- lemurs_focus_1 %>% 
  
  # calculate date missing or dead
  mutate(date_missing_dead = dob %m+% years(round(age_max_live_or_dead_y)))

# check things look like they make sense with date missing / dead calculates
ggplot(lemurs_focus_2, aes(date_missing_dead)) +
  geom_histogram()

ggplot(lemurs_focus_2, aes(dod)) +
  geom_histogram()

ggplot(lemurs_focus_2, aes(dob)) +
  geom_histogram()

lemurs_focus_2 %>% 
  filter(is.na(date_missing_dead)) %>% 
  ggplot(aes(dob)) +
  geom_histogram()
```

```{r}

known_lemurs_at_date <- function(lemur_df, date){
  
  lemur_df %>% 
    filter((dob < date) & (date <= date_missing_dead)) %>% 
    mutate(census_date = date)
}

# add dummy death dates in the future for currently alive lemurs
# to enable processing using the known_lemurs_at_date function
lemurs_focus_3 <- lemurs_focus_2 %>% 
  mutate(date_missing_dead = replace_na(date_missing_dead, ymd("2030-01-01")))

# check at the start of each calendar year which lemurs are not missing or dead
census_dates <- seq(ymd("1973-01-01"), ymd('2019-01-01'), by = 'years')

annual_lemur_census <- census_dates %>% 
  map_df(~known_lemurs_at_date(lemurs_focus_2, . )) %>% 
  mutate(age_at_census = time_length(interval(dob, census_date), "years"))

annual_lemur_census
```

```{r}
annual_lemur_census %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(y = census_date, x = age_at_census, group = census_date))

annual_lemur_census %>% 
  ggplot() +
  geom_boxplot(aes(x = census_date, y = age_at_census, group = census_date))

ggplot(annual_lemur_census) +
  ggdist::stat_gradientinterval(aes(x = census_date, y = age_at_census, group = census_date),
                                fill = "#2D1FE8", .width = 0, point_colour = NA) +
  theme_cjmr_explanatory()

annual_lemur_census %>% 
  count(census_date) %>% 
  ggplot()+
  geom_line(aes(census_date, n))


ggsave("lemur.svg", width = 250, height = 94, units = "mm")

```

```{r}
annual_lemur_census %>% 
  filter(year(census_date) %% 10 == 0) %>% 
  mutate(year = factor(year(census_date))) %>% 
  sample_frac(0.5) %>% 
  ggplot(aes(y = census_date, x = age_at_census, group = census_date)) +
  # ggdist::stat_gradientinterval(aes(x = census_date, y = age_at_census, group = census_date),
  #                               fill = "#2D1FE8", .width = 0, point_colour = NA)+
    ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.2, 
    point_colour = NA
  ) +
  geom_jitter(alpha = 0.01, height = 500) +
  #ggridges::geom_density_ridges(aes(y = census_date, x = age_at_census, group = census_date)) +
  #   geom_point(aes(y = census_date, x = age_at_census, group = census_date),
  #   ## draw horizontal lines instead of points
  #   shape = "|",
  #   size = 1,
  #   alpha = .1
  # )+
  coord_flip() +
  theme_cjmr_explanatory() + 
  facet_wrap(~year, scales = "free_x", ncol = 4)
```
How has the age profile of
```{r}

annual_lemur_census %>% 
  filter(year(census_date) %% 3 == 0) %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(y = census_date, x = age_at_census, group = census_date))

```

```{r}
age_summary <- annual_lemur_census %>% 
  mutate(year = year(census_date)) %>% 
  group_by(year, sex) %>% 
  summarise(median_age = median(age_at_census),
            oldest_age = max(age_at_census)) %>% 
  pivot_longer(cols = ends_with("age"), names_to = "age_type", values_to = "age") %>% 
  unite(col = "sex_age_type", sex, age_type) 

age_dist_by_year <- annual_lemur_census %>% 
  mutate(year = year(census_date)) 

total_num_by_year <- annual_lemur_census %>% 
  count(census_date)

total_num_by_year_sex <- annual_lemur_census %>% 
  count(census_date, sex)

prop_female <- total_num_by_year_sex %>%
  pivot_wider(names_from = sex, values_from = n) %>%
  mutate(prop_female = `F` / (`F` + `M`))

ggplot() +
  geom_area(total_num_by_year_sex, mapping = aes(census_date, n, fill = sex)) +
  geom_line(total_num_by_year, mapping = aes(census_date, n), size = 3) +
  theme_cjmr_explanatory()

# ggplot() +
#   geom_area(total_num_by_year_sex, mapping = aes(census_date, n, fill = sex), position = "fill") +
#   theme_cjmr_explanatory()

ggplot(prop_female) +
  geom_line(aes(census_date, prop_female)) +
  scale_y_continuous(limits = c(0,1)) +
  theme_cjmr_explanatory()

annual_lemur_census %>% 
  filter(year(census_date) %% 3 == 0) %>% 
  ggplot() +
  ggridges::geom_density_ridges(aes(y = census_date, x = age_at_census, group = census_date)) +
  theme_cjmr_explanatory()

ggplot() +
  geom_line(data = age_summary,
            aes(year, age, colour = sex_age_type)) +
  theme_cjmr_explanatory()
```


## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-------+
| To think about                                                                    | Notes |
+===================================================================================+=======+
| Can you think of a group of people who might view the visualisations?             |       |
+-----------------------------------------------------------------------------------+-------+
| Can you think of a specific person who might view the visualizations?             |       |
+-----------------------------------------------------------------------------------+-------+
| What do want the audience to know or be able to after viewing the visualizations? |       |
+-----------------------------------------------------------------------------------+-------+
| What expectations might the audience have about visualisations of this data?      |       |
+-----------------------------------------------------------------------------------+-------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-----------+---------------------+-----------------------------------------------+
| Question  | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+===========+=====================+===============================================+
|           |                     |                                               |
+-----------+---------------------+-----------------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Question to consider                                | Tips                                                         | Notes |
+=====================================================+==============================================================+=======+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                  |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Is the **context** explained?                       | Titles, subtitles                                            |       |
|                                                     |                                                              |       |
|                                                     | Annotate                                                     |       |
|                                                     |                                                              |       |
|                                                     | Use colour to highlight                                      |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*           |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+
| Do the **aesthetics** align with the topic/message? | Serious topic -\> dark colours, straight-forward chart types |       |
+-----------------------------------------------------+--------------------------------------------------------------+-------+

## Remember it is an iterative process
