---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files

# packages for using custom fonts
# library(systemfonts)
# library(textshaping)
library(ragg)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

Saving using custom fonts with ragg

```{r}
ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  labs(title = "22222222222222222") +
  theme(text = element_text(family = "Roboto"),
        plot.title = element_text(size = 50))

ggsave("filename.png", device = ragg::agg_png())
```


## Colour palettes

## Remember it is an iterative process

## Import data

```{r}
tuesdata <- tidytuesdayR::tt_load('2021-11-23')

writers <- tuesdata$writers
directors <- tuesdata$directors
episodes <- tuesdata$episodes
imbd <- tuesdata$imdb

```

## Clean data

```{r}
list(writers, directors, episodes, imbd) %>% 
  map(~visdat::vis_miss(.))

imbd <- imbd %>% 
  rename(episode_number = ep_num) %>% 
  mutate(first_aired = lubridate::dmy(air_date))
  
  

```

## Explore the data

```{r}
episodes %>% 
  left_join(imbd, by = c("first_aired" = "first_aired")) %>% 
  View()

  # filter(is.na(episode_number))
  # visdat::vis_miss()
  
imbd %>% 
  distinct(episode_number)

View(imbd)
View(episodes)
```


```{r}

episodes_writers <- episodes %>% 
  left_join(writers)

writer_summary <- episodes_writers %>% 
  group_by(writer) %>% 
  summarise(first_ep = min(first_aired),
            last_ep = max(first_aired),
            num_ep = n()) %>% 
  filter(num_ep > 4) %>% 
  mutate(writer_order = rank(last_ep)) %>% 
  pivot_longer(cols = c(first_ep, last_ep), names_to = "date_type", values_to = "date") %>% 
  mutate(paired = rep(1:(n()/2),each=2))
  
prolific_writers <- writer_summary %>% 
  distinct(writer) %>% 
  .$writer

pro_writers_episodes <- episodes_writers %>% 
  filter(writer %in% prolific_writers)

episode_names <- episodes %>% 
  select(episode_title, first_aired)

pro_writer_key_episodes <- writer_summary %>% 
  left_join(episode_names, by = c("date" = "first_aired"))

ggplot(mapping = aes(x = date, y = reorder(writer, writer_order))) +
  geom_line(data = writer_summary,
            mapping = aes(group = paired),
            size = 2, colour = "grey80") +
  
  geom_jitter(data = pro_writers_episodes, 
             mapping = aes(first_aired, writer, 
                           colour = uk_viewers), 
             size = 5, alpha = 0.5, height = 0.2) +
  
  scale_colour_viridis_c(option = "magma") +
  
  geom_text(data = pro_writer_key_episodes, 
            mapping = aes(date, writer, label = episode_title),
            family = "Lato") +
  
  labs(x = NULL) +
  
  theme_cjmr_explanatory()


library(ggbeeswarm)
pro_writers_episodes %>% 
  mutate(year = year(first_aired)) %>% 
  
  ggplot(mapping = aes(writer, year)) +
  geom_beeswarm(mapping = aes(colour = uk_viewers, size = rating),
                priority = "density", cex = 2) +
  
    scale_colour_viridis_c(option = "magma") +
  
  #coord_flip() +
  theme_cjmr_explanatory()
  
  #geom_jitter(width = 0)


season_averages <- episodes %>% 
  group_by(season_number) %>% 
  summarise(season_rating = mean(rating, na.rm = TRUE),
            season_mid = median(first_aired)) %>% 
  distinct(season_rating, season_mid) 

ggplot(data = episodes, 
       mapping = aes(first_aired, rating),) +
  geom_point(colour = "grey50", alpha = 0.3) +
  geom_point(data = season_averages,
             mapping = aes(season_mid, season_rating)) +
  geom_line(data = season_averages,
             mapping = aes(season_mid, season_rating)) +
  theme_cjmr_explanatory()

```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-------+
| To think about                                                                    | Notes |
+===================================================================================+=======+
| Can you think of a group of people who might view the visualisations?             |       |
+-----------------------------------------------------------------------------------+-------+
| Can you think of a specific person who might view the visualizations?             |       |
+-----------------------------------------------------------------------------------+-------+
| What do want the audience to know or be able to after viewing the visualizations? |       |
+-----------------------------------------------------------------------------------+-------+
| What expectations might the audience have about visualisations of this data?      |       |
+-----------------------------------------------------------------------------------+-------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Question to consider                                | Tips                                                         | Notes    |
+=====================================================+==============================================================+==========+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                  |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Is the **context** explained?                       | Titles, subtitles                                            |          |
|                                                     |                                                              |          |
|                                                     | Annotate                                                     |          |
|                                                     |                                                              |          |
|                                                     | Use colour to highlight                                      |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*           |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Do the **aesthetics** align with the topic/message? | Serious topic -\> dark colours, straight-forward chart types |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+

## Remember it is an iterative process

