---
title: "Tidy Tuesday: Infrastructure Investment"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading excel files
library(lubridate)  # for dates

# packages for using custom fonts
library(systemfonts)
library(textshaping)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

```{r}
mtcars %>% 
  ggplot(aes(wt, hp)) +
  geom_point() +
  theme_cjmr_explanatory()+
  #clean_axis() +
  scale_y_continuous(breaks = seq(0,350, 50), limits = c(0, 350))
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC: <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer - <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts - <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric Scherer

```{r}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space) - Cédric Scherer

```{r}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

## Colour palettes

## Remember it is an iterative process

## Import data

Note in the tidy tuesday data the chained investment amounts appear to be baselined in 2012 dollar (as standard) rather than 2021 dollar as stated in the data dictionary.

Complementary data from:

-   FRED - GDP also chained to 2012 dollars (<https://fred.stlouisfed.org/series/GDPC1>)

-   NBER - Recession data (<https://www.nber.org/research/data/us-business-cycle-expansions-and-contractions>)

```{r}
# tidy tuesday data 
tuesdata <- tidytuesdayR::tt_load('2021-08-10')
tuesdata <- tidytuesdayR::tt_load(2021, week = 33)

investment <- tuesdata$investment
chain_investment <- tuesdata$chain_investment
ipd <- tuesdata$ipd

# gdp data 
gdp <- read_xls("GDPC1.xls", skip = 10) %>% 
  janitor::clean_names()

# recessions data
recessions <- read_xlsx("bcdc_07192021_0 (1) (3).xlsx", skip = 1) %>% 
  janitor::clean_names()

```

## Clean data

Check missing data

```{r}
list(investment, chain_investment, ipd, gdp, recessions) %>% 
  map(visdat::vis_miss)



```

The recession data needs a bit more cleaning up. I am happy with the other datasets for now.

```{r}
recessions_clean <- recessions %>% 
  
  # remove rows around the actual data
  slice(-c(1:2)) %>% 
  slice(-c(35:41)) %>% 
  
  mutate(start = str_replace(peak_month_peak_quarter, "\\(.*\\)", ""),
         end = str_replace(trough_month_trough_quarter, "\\(.*\\)", ""),

        # trim trailing whitespace
        start = str_trim(start),
        end = str_trim(end),
         
         start = lubridate::my(start),
         end = lubridate::my(end) %m+% months(1)) %>% 
  
  select(start, end)

gdp_clean <- gdp %>% 
  mutate(date = date(observation_date)) %>% 
  rename(gdp = gdpc1) %>% 
  select(date, gdp) 


  
  

```

## Explore the data

I think I am going to focus on the chained investment figures

```{r}
chain_investment %>% 
  distinct(meta_cat)

chain_investment %>% 
  distinct(category)

```

```{r}
invest_overview <- chain_investment %>% 
  
  mutate(investor = !str_detect(category, "infrastructure")) %>% 
  filter(meta_cat == "Total infrastructure")
  
  ggplot(invest_overview) +
  geom_area(aes(year, gross_inv_chain, fill = category)) +
  facet_wrap(~investor) + 
  theme_cjmr_explanatory()
```

```{r}
ggplot(invest_overview) +
  geom_line(aes(year, gross_inv_chain, colour = category)) +
  facet_wrap(~investor) + 
  theme_cjmr_explanatory()
```

See page 41 of report for how infrastructure types breakdown <https://www.bea.gov/system/files/papers/BEA-WP2020-12.pdf>

```{r}
chain_investment %>% 
  filter(str_detect(str_to_lower(category), "private")) %>% 
  filter(category != "Private") %>% 
  group_by(category) %>% 
  mutate(max_invest = max(gross_inv_chain),
         high_invest = max_invest > 5e+04,
         cat_simp = case_when(
           high_invest == TRUE ~ category,
           TRUE ~ "other"
         )) %>% 
  
  ggplot() +
  geom_line(aes(year, gross_inv_chain, colour = cat_simp, group = category)) +
  theme_cjmr_explanatory()
  
```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for comparisons?**

## Think about the audience

+-----------------------------------------------------------------------------------+-------+
| To think about                                                                    | Notes |
+===================================================================================+=======+
| Can you think of a group of people who might view the visualisations?             |       |
+-----------------------------------------------------------------------------------+-------+
| Can you think of a specific person who might view the visualizations?             |       |
+-----------------------------------------------------------------------------------+-------+
| What do want the audience to know or be able to after viewing the visualizations? |       |
+-----------------------------------------------------------------------------------+-------+
| What expectations might the audience have about visualisations of this data?      |       |
+-----------------------------------------------------------------------------------+-------+

## Ask some questions to be answered by the visualizations

1.  ...

2.  ...

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

### Visualization x (question n)

+-------------+---------------------+-----------------------------------------------+
| Question    | Answer (i.e. title) | Notes to consider in design, chart types etc. |
+=============+=====================+===============================================+
|             |                     |                                               |
+-------------+---------------------+-----------------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a check-box.

+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Question to consider                                | Tips                                                         | Notes    |
+=====================================================+==============================================================+==========+
| Are the **encodings** explained?                    | Air on the side of too much explanation ...                  |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Is the **context** explained?                       | Titles, subtitles                                            |          |
|                                                     |                                                              |          |
|                                                     | Annotate                                                     |          |
|                                                     |                                                              |          |
|                                                     | Use colour to highlight                                      |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Does it have **readability** like text?             | Think visual hierarchy *(as when laying out text)*           |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+
| Do the **aesthetics** align with the topic/message? | Serious topic -\> dark colours, straight-forward chart types |          |
+-----------------------------------------------------+--------------------------------------------------------------+----------+

## Remember it is an iterative process
